---
title: "Homicides in NYC: When are they domestic?"
subtitle: ""
author: "R-Rated - Holly Ansel, Jiayi Liang, Matt Jogodnik, Kehan Zhang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height=6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(lubridate)
library(janitor)
```

```{r import}
homicide <- read_csv("data/homicide.csv")
```

```{r fix-vars}
homicide <- homicide %>%
  rename(
    perp_status = `PERP STATUS DESCRIPTION`,
    relationship = `VICTIM PERP RELATIONSHIP`,
    weapon = `WEAPON USE`,
    case_n = `CASE NUMBER`,
    record_n = `RECORD NUMBER`,
    victim_n = `VICTIM NUMBER`,
    domestic = `DOMESTIC INCIDENT`
    ) %>%
  clean_names()%>%
  mutate(
    date = mdy(date),
    month = month(date, label = TRUE),
    day = day(date),
    wday = wday(date, label = TRUE)
  )
```

## Section 1 - Introduction

For our final project, we (Team R-Rated) will be analyzing public data provided
by the New York Police Department regarding homicides in the years 2016 through
2019. Our research question is the following: 

**Which factors make a homicide more likely to be domestic/are associated with
domestic homicides?**

The homicide data we are using comes from New York City’s .gov website:
https://www1.nyc.gov/site/nypd/stats/reports-analysis/homicide.page, and is
collected by a department of the NYPD called CompStat. New York City’s police
department launched CompStat in 1994, successfully lowering crime rates through
the use of professional management, statistical analysis, and implementation.
All CompStat data is recorded by the NYPD in the Uniform Crime Reporting format.
The data is collected by CompStat because every time a homicide is committed 
and reported it goes into their records.Each entry in the data set represents 
one recorded homicide incident in New York City. The variables include the date,
the precinct, the victim and perpetrator’s age, race, and sex, along with weapon
used, arrest status, and other various notes. These variables serve to give 
important information on each of the homicides.

## Section 2 - Data analysis plan

Since our research focuses on attributes of the victim and how they are/are not
associated with domestic homicides, our main outcome variable will be domestic,
a logical (Yes/No) variable associated with whether a homicide was committed by
a family member of the victim. Predictor variables will include details
surrounding the homicide such as in what precinct it was committed in, the time
of year, and victim/perp age, sex, and race.

### Statistical Methods

We plan on using the following statistical methods:

* Statistical Hypothesis Testing

> Through the use of statistical hypothesis testing, we can see if the
proportion of homicides that are domestic in NYC is significantly different
from other populations, such as NY state or the entire United States. In order
to perform these tests, we would need domestic proportions for these areas,
however, this data should not be difficult to obtain. Furthermore, since our
data comes from multiple years, we could compare and see if one year has a
statistically significantly higher proportion of domestic homicides than another
and use this in our narrative on domestic homicides in NYC.

* Data Visualization

> Data visualization can help us break down and compare domestic homicides among
different factors like race, sex, and age. Additionally, we can use map
visualizations to show the proportion of homicides that are domestic in each
precinct to see if location has an effect on a homicide being domestic,
directly answering our research question.

* (Linear) Regression

> Along with visualizations, regression could prove to be a very helpful tool
in finding relationships between various predictor variables and homicides
being domestic (our response variable). This will allow us to determine how
several variables and the interactions between them contribute to domestic
homicide vs. non-domestic homicide.

* Classification

> Similar to linear regression, we intend to use the classification regression
methods such as using knn and logistic regression. These tools will be of use
in answering our research question based on a hypothetical homicide victim
we could predict the likelihood that they were a domestic homicide victim.

### Sample Visualizations

Below are sample visualizations that highlight potential ways in which we might
use the data to answer our research question. Underneath the code sample we 
explain the findings and how this will aid our answer of the research question.

1. Mean Age

```{r mean-age}
homicide %>% 
  summarize(mean_victim_age = mean(victim_age),
            mean_perp_age = mean(perp_age, na.rm = TRUE)
            )
```

The mean age of victims of homicide in NYC between the years 2016-2019 was
34.70 while the mean age of perpetrators was 32.88. This will be used in
general analysis of the effect age has on being a victim of domestic homicide as
well as for general hypothesis testing of age in NYC as it compares to other
areas.

2. Proportion of Homicides that are Domestic

```{r domestic}
homicide %>% 
  filter(domestic != "NA") %>% 
  count(domestic) %>% 
  mutate(
    prop_domestic = n/sum(n)
  ) %>% 
  filter(domestic == "YES") %>% 
  select(domestic, prop_domestic)
```

The proportion of homicides in NYC that were committed by relatives of the
victim between the years 2016-2019 was 23.45%. This will be used in our analysis
of domestic homicides and comparison to other areas - it's clear that a large
proportion of homicide victims in NYC thus know their killers.

3. Domestic Homicides by Race Visualization
```{r domestic-by-race}
homicide %>%
  filter(domestic != "NA" & victim_race != "NA") %>% 
  ggplot(mapping = aes(x = victim_race, fill = domestic)) +
    geom_bar(position = "fill") + 
    coord_flip() +
    labs(title = "Proportion Domestic Homicides by Race", x = "Race", 
         y = "Proportion of Domestic Homicides", fill = "Is Domestic Homicide")
```
When looking at the proportion of domestic homicides by race it is evident
that the proportion is highest among Asian/Pacific Islanders. This visualization
serves our research question to show how race can potentially contribute as a 
factor of what makes someone more likely to be a domestic homicide victim. It
also demonstrates visually that a lot of homicides are domestic.


## Section 3 - Data

See data dimensions and codebook in README.

```{r glimpse}
glimpse(homicide)
```
