---
title: "Homicides in NYC: When are they domestic?"
subtitle: "An analysis of various factors associated with domestic homicide"
author: "R-Rated - Holly Ansel, Jiayi Liang, Matt Jogodnik, Kehan Zhang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 4, fig.width = 7, fig.align = "center")
```

```{r packages}
library(janitor)
library(tidyverse)
library(lubridate)
library(broom)
library(class)
library(infer)
```

```{r import}
homicide <- read_csv("data/homicide.csv")
```

```{r fix-vars-names}
homicide <- homicide %>%
  rename(
    perp_status = `PERP STATUS DESCRIPTION`,
    relationship = `VICTIM PERP RELATIONSHIP`,
    weapon = `WEAPON USE`,
    case_n = `CASE NUMBER`,
    record_n = `RECORD NUMBER`,
    victim_n = `VICTIM NUMBER`,
    domestic = `DOMESTIC INCIDENT`
  ) %>%
  clean_names() %>%
  mutate(
    date = mdy(date),
    month = month(date, label = TRUE),
    day = day(date),
    wday = wday(date, label = TRUE)
  )
```

```{r grouping-together-vars}
homicide <- homicide %>%
  mutate(precinct = as.integer(precinct),
         weapon = if_else(weapon == "UNKNOWN", "OTHER", weapon),
         weapon = if_else(weapon == "HANDS,FISTS,ECT.", 
                          "HANDS, FISTS, FEET, ETC.", weapon),
         relationship = if_else(relationship == "OTHER-KNOWN TO VICTIM", 
                                "OTHER KNOWN TO VICTIM", relationship),
         relationship = if_else(relationship == "CHILD OF INTIMATE PARTNER" |
                                  relationship == "STEP CHILD", 
                                "STEP CHILD/CHILD OF INTIMATE PARTNER", 
                                relationship),
         relationship = if_else(relationship == "EX-INTIMATE PARTNER" |
                                  relationship == "EX-SPOUSE", 
                                "EX-SPOUSE/INTIMATE PARTNER", 
                                relationship),
         victim_race = if_else(victim_race == "ASIAN/PAC.ISL",
                               "ASIAN / PACIFIC ISLANDER",
                               victim_race)
  )
```

```{r only-domestic-data}
domestic_homicides <- homicide %>%
  filter(domestic == "YES") 
```

## Section 1 - Introduction

For our final project, we (Team R-Rated) will be analyzing public data provided
by the New York Police Department regarding homicides in the years 2016 through
2019. Our research question is the following: 

**Which factors make a homicide more likely to be domestic/are associated with
domestic homicides?**

The homicide data we are using comes from New York City’s .gov website:
https://www1.nyc.gov/site/nypd/stats/reports-analysis/homicide.page, and is
collected by a department of the NYPD called CompStat. New York City’s police
department launched CompStat in 1994, successfully lowering crime rates through
the use of professional management, statistical analysis, and implementation.
All CompStat data is recorded by the NYPD in the Uniform Crime Reporting format.
The data is collected by CompStat because every time a homicide is committed 
and reported it goes into their records. Each entry in the data set represents 
one recorded homicide incident in New York City. The variables include the date,
the precinct, the victim and perpetrator’s age, race, and sex, along with weapon
used, arrest status, and other various notes. These variables serve to give 
important information on each of the homicides.


## Section 2 - Data analysis plan

Since our research focuses on attributes of the victim and how they are/are not
associated with domestic homicides, our main outcome variable will be domestic,
a logical (Yes/No) variable associated with whether a homicide was committed by
a family member of the victim or otherwise qualifies as being domestic. 
Predictor variables will include details surrounding the homicide such as 
what precinct or borough it was committed in, the time of year, and 
victim/perp age, sex, and race. While analyzing the data, we compare the
domestic homicide with the non-domestic ones to see the role each predictable
variable plays.


### Statistical Methods

To conduct our analysis we plan on using the following statistical methods:

* Data Visualization

> Data visualization can help us break down and compare domestic homicides among
different factors like race, sex, and age. Additionally, we can use map
visualizations to show the proportion of homicides that are domestic in each
borough to see if location has an effect on a homicide being domestic,
directly answering our research question.

* Statistical Hypothesis Testing

> Through the use of statistical hypothesis testing, we can see if the
proportion of homicides that are domestic in NYC is significantly different
from other populations, such as NY state or the entire United States. In order
to perform these tests, we would need domestic proportions for these areas,
however, this data should not be difficult to obtain. Furthermore, since our
data comes from multiple years, we could compare and see if one year has a
statistically significantly higher proportion of domestic homicides than another
and use this in our narrative on domestic homicides in NYC.

* Classification

> Similar to linear regression, we intend to use logistic regression as a
method of classification. These tools will be of use in answering our research
question based on a hypothetical homicide victim we could predict the likelihood
that they were a domestic homicide victim.

### Visualizations

Below are sample visualizations that highlight simple qualities of the data that
influence our analysis for answering the research question.

The majority of our summary statistics will be introduced in section 4 to 
connect the visualizations with our deeper analyses.

1. Domestic Homicides by Race Visualization

When looking at the proportion of domestic homicides by race it is evident
that the proportion is highest among Asian/Pacific Islanders. This demonstrates
how race can potentially contribute as a factor of what makes someone more
likely to be a domestic homicide victim. This may be due to cultural factors
that potentially influence familial relations or is coincidental and due to 
other circumstances.

```{r domestic-by-race}
homicide %>%
  filter(domestic != "NA" & victim_race != "NA") %>%
  ggplot(mapping = aes(x = victim_race, fill = domestic)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("#B7D8FF", "#1A5CB5")) +
  coord_flip() +
  labs(
    title = "Proportion Homicides by Victim Race",
    subtitle = "Domestic vs Non-Domestic",
    x = "Victim Race",
    y = "Proportion of Homicides",
    fill = "Is Domestic Homicide"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
    
```


2. Domestic Homicide Probability By Borough Visualization

This visualization demonstrates the dispersion of domestic homicides across
the boroughs of New York City each year from 2016 to 2019. Upon inspection,
it is evident that Staten Island has the most domestic homicides out of the 
five boroughs. Additionally, it appears that Brooklyn has the least
domestic homicides. It is plausible that this relationship could be related
factors such as the average household income or environmental factors
such as gang presence.

```{r domestic-homicide-probability-between-boroughs, fig.height = 3}
graph_title2 <- "High Variability in Domestic Homicide Probability
Between Boroughs"

homicide %>%
  mutate(borough = factor(borough)) %>%
  group_by(borough, year) %>%
  count(domestic) %>%
  mutate(prop_domestic_per_borough_year = n / sum(n)) %>%
  filter(domestic == "YES") %>%
  ggplot(mapping = aes(x = year, y = borough,
                       fill = prop_domestic_per_borough_year)) +
  geom_tile(color = "grey") +
  scale_fill_gradient(
    low = "#EDF8FF",
    high = "#1A5CB5",
    breaks = c(0, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45)
  ) +
  theme_minimal(base_size = 8) %+replace%
  theme(legend.direction = "vertical",
        legend.key.size = unit(1, "cm")) +
  labs(
    title = str_replace_all(graph_title2, "[\r\n]" , ""),
    fill = "Probability of Domestic Homicide",
    x = "Year",
    y = "Borough"
  )
```

     
3. Proportion of Weapons Used in Domestic Homicides Compared To Regular 
Homicides Visualization and Domestic Homicide Weapons Used Summary Statistic

When examining the weapons used in domestic homicides, this visualization
indicates that the weapons used for domestic and non-domestic homicides
vary. Compared to regular homicides, domestic homicides use asphyxiation and
strangulation proportionally more, as opposed to regular homicides which use
handguns or explosives more.      
The summary statistic shows that the incidents of
neglect and narcotics, although they exclusively occurred in domestic homicides,
were infrequent. When looking more at the summary statistic of the number of
domestic homicides for each weapon type, it is evident that the most frequent
weapons in domestic homicides is cutting/stabbing, handguns, and hands, fists,
feet, etc. So although the proportions are lower compared to regular homicides,
these are the most common weapon types for domestic homicides.

```{r weapon-use}
homicide %>%
  filter(domestic != "NA" & weapon != "NA") %>%
  ggplot(mapping = aes(x = weapon, fill = domestic)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("#B7D8FF", "#1A5CB5")) +
  coord_flip() +
  labs(
    title = "Proportion of Weapons Used in Homicides",
    subtitle = "Domestic vs Non-Domestic",
    x = "Weapon",
    y = "Proportion of Homicides"
  ) +
  theme_minimal(base_size = 10)

domestic_homicides %>%
  group_by(weapon) %>%
  count() %>%
  select(weapon, n)
```

4. Distribution of the Age of Domestic Homicide Victims

The distribution of the age of domestic homicide victims illustrates that many 
victims are young children, especially infants. Among the adult population,
the victim age is relatively evenly spread.

```{r victim-age}
ggplot(data = domestic_homicides, mapping = aes(x = victim_age)) +
  geom_histogram(binwidth = 3,
                 color = "#6DB1FF",
                 alpha = 0.75) +
  labs(title = "Distribution of Age of Domestic Homicide Victims",
       x = "Victim Age", y = "Number of Domestic Homicides") +
  theme_minimal(base_size = 10)
```


5. Domestic Homicides by Sex Visualization

In looking at the proportion of homicides by sex and distinguishing between
domestic and non-domestic homicides it is evident that the proportion
of females who are victims of domestic homicides is much higher than that of
males.

```{r domestic-by-sex}
homicide %>%
  filter(domestic != "NA" &
           victim_sex != "NA" & victim_sex != "U") %>%
  ggplot(mapping = aes(x = victim_sex, fill = domestic)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("#B7D8FF", "#1A5CB5")) +
  labs(
    title = "Proportion Homicides by Victim Sex",
    subtitle = "Domestic vs Non-Domestic",
    x = "Victim Sex",
    y = "Proportion of Homicides",
    fill = "Is Domestic Homicide"
  ) +
  theme_minimal(base_size = 10)
```

## Section 3 - Data

See data dimensions and code book in README.

```{r glimpse}
glimpse(homicide)
```

## Section 4 - Methods and Results

Before we do some further data analysis, we will first set the seed so our
results are reproducible.

```{r seed}
set.seed(71189752)
```

Next, we have to make some adjustments to the data-set to clean up our data.
Since we will be running further tests on age, sex, and race (and obviously
their relationship to domestic homicide), these are the 4 variables we will
keep. Additionally, since we will be doing hypothesis testing, we select a 
small subset of our population (~20% = 200) so that our results can be
compared with the entire population.

```{r ht-setup}
homicide_ht <- homicide %>%
  select(victim_age, victim_sex, victim_race, domestic) %>%
  na.omit() %>%
  mutate(
    victim_age_group = case_when(
      victim_age < 10 ~ "<10",
      victim_age >= 10 & victim_age < 20 ~ "10-19",
      victim_age >= 20 & victim_age < 30 ~ "20-29",
      victim_age >= 30 & victim_age < 40 ~ "30-39",
      victim_age >= 40 & victim_age < 50 ~ "40-59",
      victim_age >= 50 & victim_age < 60 ~ "50-59",
      victim_age >= 60 ~ ">59"
    )
  )

indices <- sample(nrow(homicide_ht), 200)

homicide_ht_sample <- homicide_ht %>%
  slice(indices)
```


## Proportion of Homicides that are Domestic

Using the simulation based method on a smaller sample of the data, we can 
generate confidence intervals for the proportion of homicides that are domestic.
This statistic does a good job summarizing a key aspect of the data set that we
are working with.

```{r simulation-based-confidence-interval}
set.seed(71189752)

domestic_dist <- homicide_ht_sample %>%
  specify(response = domestic, success = "YES") %>%
  generate(domestic_data, reps = 5000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  summarize(lower_bound = quantile(stat, 0.025),
            upper_bound = quantile(stat, 0.975))

domestic_dist
```

This confidence interval is generated with a simulation-based method, and
indicates that we are 95% confident that the interval from 0.17 to 0.285
captures the true population proportion of homicides that are domestic in NYC
from the years 2016 through 2019.

```{r domestic}
homicide %>%
  filter(domestic != "NA") %>%
  count(domestic) %>%
  mutate(prop_domestic = n / sum(n)) %>%
  filter(domestic == "YES") %>%
  select(domestic, prop_domestic)
```

As we can see by this sample statistic that looks at the entire data set, both
statistics are extremely similar for the proportion of domestic homicides. The
proportion of homicides in NYC that were committed by relatives of the victim or
otherwise qualify as domestic between the years 2016-2019 was actually 23.45%,
which lies within the confidence interval. This summary statistic demonstrates
domestic homicides are not a rare event and that frequently homicide victims in
NYC know and are very close to their killers. The similarity between our
confidence interval and the summary statistic demonstrates that this sample
which will be used in later parts of our analysis is representative of the 
population. 


## The Influence of Victim's Age

As seen in our visualization from part 2, age of victims of domestic homicide
seems to be skewed very far to the left, suggesting people at very low age are
disproportionately affected by domestic homicide. The summary table below 
confirms this, with approximately 92.7% of homicides involving a victim under 
10 being domestic.

```{r summary-age}
homicide_ht %>%
  count(victim_age_group, domestic) %>%
  group_by(victim_age_group) %>%
  mutate(prop = n / sum(n),
         total = sum(n)) %>%
  filter(domestic == "YES") %>%
  arrange(desc(prop)) %>%
  select(victim_age_group, prop, total)
```

With this in mind, we wanted to examine the influence of victims' age on
their likelihood of being a homicide victim so we performed an stimulation-based
hypothesis test on the correlation between age and whether a homicide is
domestic. For all the hypothesis tests in the project, we adopted the standard
$\alpha = 0.05$.     

Hypotheses are as follows:

$H_0:$ Victim age and whether a homicide is domestic are independent.

$$p_{<10} = p_{\geq10}$$

$H_1:$ Victim age and whether a homicide is domestic are not independent
      (in fact, people under the age of 10 are more likely to be the victim for
      domestic homicide).
    
$$p_{<10} > p_{\geq10}$$

```{r ht-age}
set.seed(71189752)

p_hat_diff_age <- homicide_ht_sample %>%
  mutate(victim_under_10 = if_else(victim_age_group == "<10", "Y", "N")) %>%
  count(victim_under_10, domestic) %>%
  group_by(domestic) %>%
  mutate(p_hat = n / sum(n)) %>%
  filter(victim_under_10 == "Y") %>%
  pull(p_hat) %>%
  diff()

null_age_dist <- homicide_ht_sample %>%
  mutate(victim_under_10 = if_else(victim_age_group == "<10", "Y", "N")) %>%
  specify(response = domestic,
          explanatory = victim_under_10,
          success = "YES") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 2000, type = "permute") %>%
  calculate(stat = "diff in props", order = c("Y", "N"))

visualize(null_age_dist) +
  shade_p_value(
    obs_stat = p_hat_diff_age,
    direction = "greater",
    color = "#1A5CB5",
    fill = "#B7D8FF"
  )+
  labs(x = "difference in proportions of victim under 10")

null_age_dist %>%
  get_p_value(obs_stat = p_hat_diff_age, direction = "greater")
```

We find the p-value to be 0.029, which is less than the standard alpha of 0.05.
Therefore, we can reject our null hypothesis that victim age and domestic
homicide are independent. Extrapolated to the population as a whole, this poses
the possibility (though does not definitively state) that victim age and
the likelihood of a homicide being domestic could be linked.

## The Influence of Victim's Race

Furthermore, our visualization from part 2 shows that the race of victims of
domestic homicide seems to be linked with a homicide being domestic, with
Asian/Pacific Islanders showing the highest proportion of domestic homicide.
The summary table below confirms this, with approximately 34% of homicides
involving an Asian/Pacific Islander victim being domestic.

```{r summary-race}
homicide_ht %>%
  count(victim_race, domestic) %>%
  group_by(victim_race) %>%
  mutate(prop = n / sum(n),
         total = sum(n)) %>%
  filter(domestic == "YES" & victim_race != "UNKNOWN") %>%
  arrange(desc(prop)) %>%
  select(victim_race, prop, total)
```

Seeing the proportions of different races as domestic homicide victims led
us to perform a hypothesis test with the following hypotheses:

$H_0:$ Victim race and whether a homicide is domestic are independent.

$$p_{victim = asian/pi} = p_{victim \neq asian/pi} $$

$H_1:$ Victim race and whether a homicide is domestic are not independent
      (in fact, the proportion of domestic homicides among Asian/Pacific
      Islander victims is greater than among other races).
    
$$p_{victim = asian/pi} > p_{victim \neq asian/pi} $$

```{r ht-race}
set.seed(71189752)

p_hat_diff_race <- homicide_ht_sample %>%
  mutate(victim_asianpi = if_else(victim_race == "ASIAN / PACIFIC ISLANDER",
                                  "Y", "N")) %>%
  count(victim_asianpi, domestic) %>%
  group_by(domestic) %>%
  mutate(p_hat = n / sum(n)) %>%
  filter(victim_asianpi == "Y") %>%
  pull(p_hat) %>%
  diff()

null_race_dist <- homicide_ht_sample %>%
  mutate(victim_asianpi = if_else(victim_race == "ASIAN / PACIFIC ISLANDER",
                                  "Y", "N")) %>%
  specify(response = domestic,
          explanatory = victim_asianpi,
          success = "YES") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 2000, type = "permute") %>%
  calculate(stat = "diff in props", order = c("Y", "N"))

visualize(null_race_dist) +
  shade_p_value(
    obs_stat = p_hat_diff_race,
    direction = "greater",
    color = "#1A5CB5",
    fill = "#B7D8FF"
  ) +
  labs(x = "difference in proportions of victim as Asian/pacific islander")

null_race_dist %>%
  get_p_value(obs_stat = p_hat_diff_race, direction = "greater")
```

We find the p-value to be 0.38, which is far greater than a standard alpha
of 0.05. So we fail to reject our null hypothesis that victim race and
domestic homicide are independent because we have insufficient evidence.

## The Influence of Victim's Sex

Additionally the visualization relating to the sex of victims of domestic 
homicide from section 2 shows that female has a higher possibility of being the
victim of domestic homicide. The summary table below confirms this, with
approximately 64.5% of homicides involving a female being domestic compared to
only 12.3% involving males being domestic.

```{r summary-sex}
homicide_ht %>%
  count(victim_sex, domestic) %>%
  group_by(victim_sex) %>%
  mutate(prop = n / sum(n),
         total = sum(n)) %>%
  filter(domestic == "YES" & victim_sex != "U") %>%
  arrange(desc(prop)) %>%
  select(victim_sex, prop, total)
```

Similar to the previous variables analyzed, this led us to perform a hypothesis
test with the following hypotheses:

$H_0:$ Victim sex and whether a homicide is domestic are independent.

$$p_{female} = p_{male} $$

$H_1:$ Victim sex and whether a homicide is domestic are not independent
      (in fact, the proportion of domestic homicides among female victims is
      greater than among male victims).
    
$$p_{female} > p_{male} $$

```{r ht-sex}
set.seed(71189752)

p_hat_diff_sex <- homicide_ht_sample %>%
  mutate(victim_female = if_else(victim_sex == "F", "Y", "N")) %>%
  count(victim_female, domestic) %>%
  group_by(domestic) %>%
  mutate(p_hat = n / sum(n)) %>%
  filter(victim_female == "Y") %>%
  pull(p_hat) %>%
  diff()

null_sex_dist <- homicide_ht_sample %>%
  mutate(victim_female = if_else(victim_sex == "F", "Y", "N")) %>%
  specify(response = domestic,
          explanatory = victim_female,
          success = "YES") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 2000, type = "permute") %>%
  calculate(stat = "diff in props", order = c("Y", "N"))

visualize(null_sex_dist) +
  shade_p_value(
    obs_stat = p_hat_diff_sex,
    direction = "greater",
    color = "#1A5CB5",
    fill = "#B7D8FF"
  ) +
  labs(x = "difference in proportions of female victim")

null_sex_dist %>%
  get_p_value(obs_stat = p_hat_diff_sex, direction = "greater")
```

We find the p-value to be 0, which being less than a standard alpha of 0.05
means that we can reject our null hypothesis that victim sex and domestic
homicide are independent. Extrapolated to the population as a whole, this poses
the possibility (though does not definitively state) that victim sex and
the likelihood of a homicide being domestic could be linked.

## Logistic Regression

Furthering our visualizations from part 2 on the effects borough, weapon, age,
and sex have on the likelihood of a homicide being domestic, we will now perform
logistic regression in an attempt to classify some observations in our data-set.

```{r logit}
set.seed(71189752)

homicide_lm <- homicide %>%
  select(
    month,
    wday,
    day,
    year,
    weapon,
    victim_age,
    victim_sex,
    victim_race,
    perp_age,
    perp_sex,
    perp_race,
    domestic,
    precinct,
    borough
  ) %>%
  na.omit() %>%
  mutate(
    month = factor(month),
    wday = factor(wday),
    day = factor(day),
    year = factor(year),
    weapon = factor(weapon),
    victim_sex = factor(victim_sex),
    victim_race = factor(victim_race),
    perp_sex = factor(perp_sex),
    perp_race = factor(perp_race),
    precinct = factor(precinct),
    borough = factor(borough),
    domestic = if_else(domestic == "YES", 1, 0)
  )

indices <- sample(nrow(homicide_lm), 50)

homicide_test <- homicide_lm %>%
  select(-domestic) %>%
  slice(indices)

homicide_train <- homicide_lm %>%
  select(-domestic) %>%
  slice(-indices)

train_status <- homicide_lm %>%
  select(domestic) %>%
  slice(-indices) %>%
  pull()

true_status <- homicide_lm %>%
  select(domestic) %>%
  slice(indices) %>%
  pull()

homicide_logit <- homicide_lm %>%
  slice(-indices)

logit_domestic <-
  glm(
    domestic ~ weapon + victim_age + victim_sex + borough,
    data = homicide_logit,
    family = "binomial"
  )

logit_domestic2 <-
  glm(domestic ~ victim_age + victim_sex + borough,
      data = homicide_logit,
      family = "binomial")

logit_domestic3 <- glm(domestic ~ victim_age + victim_sex,
                       data = homicide_logit,
                       family = "binomial")

logit_domestic4 <-
  glm(
    domestic ~ weapon + victim_age + victim_sex + borough +
      month,
    data = homicide_logit,
    family = "binomial"
  )

logit_domestic %>%
  tidy()

logit_domestic %>%
  glance() %>%
  pull(AIC)

logit_domestic2 %>%
  tidy()

logit_domestic2 %>%
  glance() %>%
  pull(AIC)

logit_domestic3 %>%
  tidy()

logit_domestic3 %>%
  glance() %>%
  pull(AIC)

logit_domestic4 %>%
  tidy()

logit_domestic4 %>%
  glance() %>%
  pull(AIC)
```

We test 4 logistic regression models, using calculated AIC values to determine
the best model. We find the first to have the lowest AIC, and thus choose this
one for classification purposes.

```{r log-class}
set.seed(71189752)

pred_probs <- augment(logit_domestic, newdata = homicide_test) %>%
  mutate(pred_prob = exp(.fitted) / (1 + exp(.fitted)))
pred_results <- pred_probs %>%
  mutate(type = case_when(pred_prob < 0.5 ~ 0, pred_prob >= 0.5 ~ 1)) %>%
  pull(type)
mean(pred_results == true_status)
```

We find that, given our first model, 78% of our test data was classified
correctly. This figure suggests we have built a successful logistic regression
model.

And, just to confirm that race is not a great factor for determining domestic
homicide, we will incorporate it with our best model.

```{r log-race}
set.seed(71189752)

logit_race <-
  glm(
    domestic ~ weapon + victim_age + victim_sex + borough +
      victim_race,
    data = homicide_logit,
    family = "binomial"
  )

logit_race %>%
  glance() %>%
  pull(AIC)
```

We find that AIC has increased by adding race, and thus, adding race worsens the
fit of the model, agreeing with our hypothesis test above.

## Section 5 - Discussion

For our final project, we focused on determining the relationship between
various factors of homicide and the likelihood of a homicide being domestic. We
conducted basic hypothesis tests of independence, bootstrapping intervals of
confidence, and logistic regression to determine which of our data set’s factors
best predicted or were connected to domestic homicides. We conclude that
domestic homicides are in fact influenced by many factors, among them and in
loose proposed order of effect victim sex, victim age, weapon used, and borough.
The extent to which these factors affect domestic homicide is subject for
further analysis.

We started our analysis by conducting several preliminary visualizations which
led us to select victim age, victim sex, victim race, weapon used, and borough
as our factors to explore. Hypothesis testing for independence conducted on
victim age and victim sex rejected our null hypotheses, suggesting a possible
connection between these factors and domestic homicide. For victim sex, our
analysis showed that females are much more likely to be domestic homicide
victims than males. Similarly for age our hypothesis test concluded that 
young children (under the age of ten) are the most susceptible age group
to be a victim of domestic homicide. Conversely, we failed to reject our null
hypothesis for race, so we have insufficient evidence to make a conclusion about
the role of race in domestic homicide victims. Logistic regression also
supported a connection between domestic homicide and victim age and sex as the
best model found incorporate these factors with decent (78%) success. Logistic
regression also suggested a connection between borough and weapon used and
domestic homicide, suggesting these factors may also be good predictors of
domestic homicide. Logistic regression incorporating victim race increased its
AIC, which was in agreement with results from the hypothesis test suggesting
race to be a lesser factor in determining likelihood of domestic homicide.

In tying these conclusions back to our research question our current analysis
presents females and children as two groups that are more likely to be victims
of domestic homicide. Furthermore through the logistic regression we have also
seen that the borough one lives in, such as Staten Island, may influence the
likelihood that someone is a domestic homicide victim. However, our analysis
indicated that we do not have enough evidence to make conclusions about the 
effect of the victim's race. The logistic regression also demonstrated that
specific methods or weapons can serve as predictors for domestic homicide
so the form of the homicide differs compared to that of a stranger. Together
the analysis shows that the circumstances revolving around domestic homicides
differ from those of regular homicides, and unfortunately, are not uncommon 
events in NYC.

Reflecting upon our methods, it would’ve been helpful to compare logistic 
regression classification results with those from k-NN, however, the fact that 
our data set was almost entirely comprised of categorical variables made this 
impossible as k-NN relies on distance. This also prevented us from developing 
linear models and using the step function to arrive at the best one – our 
original plan for classification modeling. Additionally, a chi-square test for 
independence could have tested all levels of a factor rather than just 
selecting the highest proportion level. However, we had difficulty in 
interpreting these results and thus stuck with our original analysis. This 
would have nonetheless made our hypothesis tests more reliable for what we were 
testing. Finally, it could have helped our analysis to make demographic 
comparisons to the complete NYC population. It is possible that disparities in 
domestic homicide proportion are high among certain groups simply because there 
are fewer observations and thus the data set isn’t representative of this 
demographic. This is supported by both Asian/Pacific Islander and Victim <10 
having only around 50 observations each. Further analysis of domestic homicide 
in NYC could address these concerns to arrive at a more detailed and probable 
conclusion.

Further analysis on this data set should focus on improving the reliability and 
level of detail of the conclusion. This would likely include starting out by 
creating more summary statistics to create more numerical variables which would 
allow for greater modeling/bootstrapping. Analyzing and investigating more
variables would also create greater confidence in those selected. Comparisons
along the way with actual population statistics from NYC would help put results
and proportions in context with their demographic’s relative representation. 
And analysis of multiple variables simultaneously such as seeing the difference
between male and female children under the age of 10 as homicide victims would
also give a deeper understanding. Additionally doing more analysis on whether
there are multiple victims in one domestic homicide incident.Finally, chi-square
testing would allow for better analysis of the relationship between each
variable’s several factor levels and domestic homicide. These changes would 
create more robust and generalizable results, leading to a better analysis of 
our research question: which factors of homicide increase its likelihood of 
being domestic.

Additionally if we were to start over with this project it might have been 
useful to also analyze more the relationships between the victim and 
perpetrator in domestic homicides. Doing more analysis on the perpetrator 
would give us a bigger picture view on domestic homicides because it is not 
random but rather there is a close relationship between the perpetrator and
victim. This would allow deeper analysis on elements such as if fathers are 
more often the perpetrator. By looking more at this side of the question, 
our conclusions would contain a broader view of the circumstances relating
to domestic homicides, not just that of the victim.

